My Data Literacy Project\\ (Replace this with your Project Title)

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
update result

% In this section, describe \emph{what you did}. Roughly speaking, 
% explain what data you worked with, how or from where it was collected, it's structure and size.
%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
% If you downloaded data from the net, show an exploratory analysis that builds
% intuition for the data, and shows that you know the data well. 
% If you are doing a custom analysis, explain how it works and why it is the right choice.
% If you are using a standard tool, it may still help to briefly outline it. Cite relevant works. 
% You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.
\section{Data and Methods}\label{sec:methods}
In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}., which contains the
total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
for a study in Tübingen.
The second dataset is from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} which contains data about the high-school graduates from 1970 to 2023.
Our third and last dataset is from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} and contains the salaries of the large industrial sectors in Germany from 2005 to 2021. 
One main problem with the mapping between university subjects and actual jobs is that there exists no perfect one-to-one mapping. 
We also removed the inflation from the salaries to get the real wage, because the inflation could 
cause a false pattern detection in our autoregressive model. 
To reduce the amount of bias introduced in this step, we also used the Inflation data from the Deutsches Statistisches Bundesamt over the years.\\

@@

\begin{figure*}
    \includegraphics{fig/prediction_lag_1.pdf}
    \caption{Prediction based on lag order 1}
\end{figure*}
As you can see, the prediction of the average salary after tax is almost linear, but the true average salary after tax has a totally different shape. After further analysis, we decided to use the hyperparameter p based
on an information criterion namely AIC and BIC. The abbreviation AIC stands for Akaike information criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting. 

@@

The optimal lag based on AIC is 5, while for BIC it is 1, which aligns with the PACF.
\begin{figure*}
    \includegraphics{fig/prediction_lag_5.pdf}
    \caption{Prediction based on lag order 5}
\end{figure*}
As you can see the lag order 5 mimics the average salary after tax better than the lag order 1.\\

@@

\section{Results}\label{sec:results}

In this section outline your results. At this point, you are just stating the outcome of your analysis. 
You can highlight important aspects (``we observe a significantly higher value of $x$ over $y$''), but leave interpretation and opinion to the next section. This section absoultely \emph{has} to include at least two figures.

\begin{figure*}
    \includegraphics{fig/params.pdf}
    \caption{This is a figure caption. Write this, if we end up using that fig ;-)}
\end{figure*}

\section{Discussion \& Conclusion}\label{sec:conclusion}

Use this section to briefly summarize the entire text. Highlight limitations and problems, but also make clear statements where they are possible and supported by the analysis.

>>>>

% In this section, describe \emph{what you did}. Roughly speaking, 
% explain what data you worked with, how or from where it was collected, it's structure and size.
%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
% If you downloaded data from the net, show an exploratory analysis that builds
% intuition for the data, and shows that you know the data well. 
% If you are doing a custom analysis, explain how it works and why it is the right choice.
% If you are using a standard tool, it may still help to briefly outline it. Cite relevant works. 
% You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.

\section{Data and Methods}\label{sec:methods}
In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}., which contains the
total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
for a study in Tübingen.
The second dataset is from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} which contains data about the high-school graduates from 1970 to 2023.
Our third and last dataset is from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} and contains the salaries of the large industrial sectors in Germany from 2005 to 2021. 
One main problem with the mapping between university subjects and actual jobs is that there exists no perfect one-to-one mapping. 
We also removed the inflation from the salaries to get the real wage, because the inflation could 
cause a false pattern detection in our autoregressive model. 
To reduce the amount of bias introduced in this step, we also used the Inflation data from the Deutsches Statistisches Bundesamt over the years.\\

@@

\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 1.25723662 & 1.27163776 & 3.94417629 \\
        \hline
        Prediciton 2 & 4.84256257 & 0.11922267 & 1.72230067
    \end{tabular}
    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 1}
\end{table}
As you can see, the percent difference between the true values especially for the students are quite high, becasue of that we tried the AIC and BIC to find a better lag for the 
hyperparameter p.  The abbreviation AIC stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting. 

@@

The optimal lag based on AIC is 6, while for BIC it is 1, which aligns with the PACF.
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 0.48413863 & 0.41885829 & 0.11117345 \\
        \hline
        Prediciton 2 & 1.34596809 & 0.30583828 & 0.93859448
    \end{tabular}
    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 5}
\end{table}
As you can see there is a major improvement in the prediction for the students in comparison to lag order 1. We decided to use lag 5, because the main focus of our 
of our research is to predict the students based on salary and graduates.

@@

\section{Results}\label{sec:results}
The hyperparameter optimization is always a trade-off between several optimal values for different parameters. 
Our main goal was to predict the students who enroll in study subjects in the STEM sector when using the salary and the
graduates of the given years as prior knowledge. For this reason, we used a lag of 5, because the overall percentage error
was the lowest for the students.
\begin{figure}[ht]
    \includegraphics{fig/prediction_5.pdf}
    \caption{VAR Model Prediction with hyperparameter p=5}
\end{figure}
During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
of predicted students should go down, because fewer students imply that there is a lack of qualified workers on the job market.
This on the other hand leads to an increase in the salary. Now we manipulated the salary data with a fixed $\alpha\in \mathbb{N}$ to increase
the whole time series by a constant factor. We used this approach because the vector autoregressive model internally uses Mean Squared Error as loss function.
The Mean Squared Error squares the difference between the data points, which leads to a stronger influence of large differences.
This high error leads to a wrong minimization, which gives wrong coefficients.
Another more important point is that we don't want to violate stationarity, which ensures that our model predicts in a good manner.\\
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates\\
        \hline
        students & 1.000000 & -0.115442 &  0.111069\\
        \hline
        salary & -0.115442 & 1.000000 & -0.902321\\
        \hline
        graduates & 0.111069 & -0.902321 &  1.000000\\
    \end{tabular}
    \caption{Correlation Matrix of Var(5) model}
\end{table}

As you can see the correlation matrix, tells us that the model also takes the relationship into account, that a 
manipulated salary has an inverse relationship to the amount of students.

\begin{figure}[H]
    \centering
    \includegraphics{fig/prediction_modified_10\%_increase_lags_5.pdf}
    \caption{Increased salary by 10 \% in Var(5) model}
\end{figure}

% Use this section to briefly summarize the entire text. Highlight limitations and problems, 
% but also make clear statements where they are possible and supported by the analysis. 
\section{Discussion \& Conclusion}\label{sec:conclusion}
